Ectopic hAMH-driven SOX17 expression induces hyperplastic Sertoli valve formation in mouse testes

Version 1

Abstract

In the terminal segment of the seminiferous tubules, SOX17 expression in the rete testis (RT) epithelium plays a crucial role in the formation of the Sertoli valve (SV), as revealed by phenotypic analyses of RT-specific Sox17 conditional knockout (cKO) mouse testes. In these RT-specific Sox17 cKO testes, SV disruption leads to the backflow of RT fluid into the seminiferous tubules, resulting in defective spermiogenesis and male infertility. Although valve deformation in the Sox17 cKO testes is likely caused indirectly by impaired downstream actions of Sox17 in the RT, the mechanisms by which SOX17 in RT influences SV formation in the seminiferous tubules remain unclear. To address this, we generated a novel AMH-Sox17 transgenic (Tg) mouse line carrying a human AMH promoter-driven Sox17 cDNA cassette. We analyzed the phenotypes of the Sertoli valve and spermatogenesis in AMH-Sox17 Tg mice, as well as in RT-specific Sox17 cKO; AMH-Sox17 Tg double mutant mice. Ectopic SOX17 (SOX17+) expression in Sertoli cells resulted in excessive Sertoli valve structures with acetylated tubulin bundles in the terminal segment of the AMH-Sox17 Tg testes, along with enhanced WNT4/RSPO1 signaling, suggesting the enhanced valve formation of ectopic SOX17+ Sertoli cells by themselves. Moreover, the AMH-Sox17 Tg could partially rescue the SV deformation and infertility in RT-specific Sox17 cKO mice, leading to proper SV formation, normal spermiogenesis and a partial recovery of male fertility in AMH-Sox17 Tg; RT-specific Sox17 cKO double mutant mice. These findings genetically demonstrate that ectopic SOX17+ Sertoli cells can compensate for SOX17 paracrine signaling in the RT, underscoring a key shared downstream pathway between RT and SV.

Introduction

Valves are critical anatomical structures that ensure unidirectional fluid flow in the circulatory and lymphatic systems. They function passively to prevent backflow, maintaining efficient fluid dynamics within vessels (Moore and Bertram, 2018; O’Donnell and Yutzey, 2020). When valve integrity is compromised, it distorts the luminal architecture, reduces intravascular pressure, and may result in regurgitation.

In the mammalian testis —an organ characterized by active, oscillatory fluid flow (Fleck et al., 2021)— a valve-like structure called the Sertoli valve (SV) emerges at the termini of seminiferous tubules (STs) during puberty (Aiyama et al., 2015; Nagasawa et al., 2018; Imura-Kishi et al., 2021). The SV comprises a specialized subset of Sertoli cells that exhibit elevated levels of phosphorylated AKT and STAT3. These cells extend cytoplasmic projections enriched with acetylated tubulin-positive (ace-TUB+) microtubules into the rete testis (RT) lumen, forming a dynamic and structurally distinct valve-like barrier (Russell, 1993). Recently, we identified SOX17, a transcription factor expressed in RT epithelial cells, as essential for SV formation and for spermiogenesis in the ST. In mice with RT-specific conditional knockout of Sox17 (Sox17 cKO), the SV fails to form properly before puberty, allowing retrograde fluid flow from the RT into the convoluted ST (Uchida et al., 2022). These findings imply the critical role of the RT-SV structure in maintaining proper fluid dynamics essential for spermatogenesis.

Interestingly, the SV epithelium provides a specialized niche for undifferentiated spermatogonia (Aundiff) and immature/proliferative Sertoli cells (Aiyama et al., 2015; see reviews by Figueiredo et al., 2021; Seco-Rovira et al., 2025). The basal compartment of the SV epithelium supports self-renewal of Aundiff spermatogonia by maintaining high expression of growth factors such as FGFs (Takashima et al., 2015; Kitadate et al., 2019; Yang et al., 2021; also see the review by Hofmann and McBeath, 2022) and GDNF (Meng et al., 2000; Uchida et al., 2016), together with by locally suppressing retinoic acid (RA)-induced differentiation (Imura-Kishi et al., 2021). Furthermore, SV Sertoli cells display Cyclin D1-positive states in mice and hamsters, and in rats and hamsters, some of these cells proliferate and contribute to the convoluted ST even at adult stage (Aiyama et al., 2015; Figueiredo et al., 2019), suggesting that the SV acts as a niche for proliferative Sertoli cells in hamsters and rats. Specification of these SV Sertoli cells is thought to be regulated by paracrine signals such as RSPO1, FGFs etc. from adjacent SOX17-positive RT cells (Imura-Kishi et al., 2021; Uchida et al., 2022; see review by Han et al., 2026). However, how SOX17-positive RT cells regulate SV formation in the proximal region of rodent testes remains unclear.

To address this question, we generated a novel transgenic (Tg) mouse line that exhibits excess SV formation by the Sertoli cells ectopically expressing Sox17 under the control of the human AMH promoter (SOX17+ Sertoli cells). Such SOX17+ Sertoli cells enhance the expression of WNT4 and RSPO1 in SV and RT regions, respectively. Moreover, we showed a partial rescue by the SOX17+ Sertoli cells in the Sertoli valve formation and aberrant spermiogenesis in the cKO testes with Sox17-null RT. These findings highlight the importance of SOX17-downstream paracrine actions in the SV and RT for the proper valve formation, rather than the SOX17 roles in the RT formation in itself.

Results

Overexpression of SOX17 in Sertoli cells induced hyperplasia of the SV region

To investigate the downstream paracrine effects of Sox17, we generated AMH-Sox17 transgenic (Tg) mice that ectopically express Sox17 in Sertoli cells under the control of the human anti-Müllerian hormone (AMH) promoter (Fig. 1A, S1). Among three Tg founder lines, one male (#27) exhibited infertility and was smaller in size of testes compared to his wild-type littermates. In this Tg male, the Sertoli valve (SV) structure appeared to be expanded in the terminal segments of the seminiferous tubules (Fig. 1B, C). Additionally, abnormal aggregates of Sertoli cells were frequently observed on the luminal side of some seminiferous tubules (indicated by asterisks in Fig. 1B). Immunostaining for SOX17 and E-cadherin in the #27 Tg testis revealed ectopically high levels of SOX17 expression in Sertoli cells throughout both the convoluted seminiferous tubules and the SV-like regions, similar to those observed in the RT epithelium (Fig. 1D). This ectopic, high-level expression of SOX17 in Sertoli cells appears to disrupt their normal function across the entire testis, ultimately leading to infertility in the #27 Tg male.

Hyperplastic SV structure in the AMH-Sox17 Tg male mouse (F0)

(A) Schematic diagram of the DNA construct containing Sox17 coding sequences driven by the human AMH promoter. (B, C) Gross morphology and hematoxylin and eosin (H&E)-stained sections of testes from infertile Tg #27 and its wildtype littermates. In the Tg testis, normal spermatogenesis is partially observed in the convoluted seminiferous tubules (ST), while an expanded SV-like region including decidual cell aggregates is evident in the proximal area near the rete testis (RT) (asterisks in C). (D) Immunostaining for SOX17 and E-cadherin (E-cad; a marker for both RT epithelia and Aundiffer spermatogonia), showing ST regions (left two panels) and SV regions (rightmost panel) of the testis sections. The AMH-Sox17 Tg testis exhibits ectopic high expression of SOX17 in almost all of the Sertoli cells in both ST and SV. Asterisks in (D) indicate protruded SOX17-positive SV Sertoli cells within the E-cad-positive RT lumen. RT, rete testis; SV, Sertoli valve. Scale bars: 1 mm (B), 200 μm (C), 50 μm (D).

Establishment of the AMH-Sox17 Tg Line (#26) with Excess SV Formation

One Tg male (#26) was fertile, and the AMH-Sox17 transgene was successfully transmitted to its offspring (Fig. S1A, B). The morphology and weight of the testes in Tg mice were comparable to those of their wild-type littermates. Although the AMH-Sox17 transgene was moderately activated in a subset of Sertoli cells during both fetal and postnatal stages (Fig. S1C, D), an expansion of the acetylated tubulin-positive Sertoli valve (SV) region was observed in the proximal testis of this line (Fig. 2A), similar to that seen in the infertile #27 Tg male with a more severe phenotype (Fig. 1D). Because this Tg line could be stably maintained by backcrossing beyond the 8th generation, line #26 was used for all subsequent experiments (hereafter referred to as “Tg”).

Morphological characterization of the hyperplastic SV structure in a novel AMH-Sox17 Tg line

(A) Immunofluorescence staining for acetylated tubulin (red) and cyclin D1 (green) in the testes from Tg (#26) mice, showing expanded SV Sertoli cells. (B) Transmission electron microscopy (TEM) analysis of a Tg (#26) testis at the RT-SV boundary (red arrow), revealing abnormal features of SV Sertoli cells, such as protruded SV (pSV) structures extending into the RT lumen (upper left in B), multiple layers of cell processes containing microtubule bundles (blue asterisks), and open extracellular spaces in the basal region of the RT-SV boundary (red asterisk). Lower panels (B’–B’’’) are magnified views of the upper right panel in B. A vertical dashed line at the center marks the region hidden by the grid (grid spacing is 10 μm). BL, multilayered SV-typical basal lamina; RT, rete testis; SV, Sertoli valve. Scale bars: 50 μm (A), 10 μm (B), 2 μm (B’–B’’’).

In this Tg line, no appreciable alterations in the RT-SV region are observed during the immature stages up to 1 week after birth. However, following sexual maturation, immunostaining for acetylated tubulin and Cyclin D1 clearly marks an expanded SV region at the terminal segments of the seminiferous tubules (Fig. 2A). This expansion was further confirmed by electron microscopy, which revealed multiple cytoplasmic layers containing bundled microtubules extending distally from Sertoli cells into the rete testis (RT) lumen (blue asterisks, Fig. 2B). At the boundary between the SV and RT regions, SV Sertoli cells were supported by a multilayered basal lamina, consistent with previous reports (Fig. S2A; Tainosho et al., 2011; Uchida et al., 2022). Notably, in the Tg testes, the SV Sertoli cells displayed vacuole-like intercellular spaces in the basal portion of the terminal segment, and some appeared detached from the basal lamina (red asterisks, Fig. 2B). Furthermore, these cells were often found protruding into the RT lumen (designated as “pSV” in Fig. 2A, 2B, S2B), where necrotic or damaged cells were observed at the distal tips of the protrusions. These findings suggest that the protruding SV Sertoli cells may undergo necrosis and be sloughed off into the RT lumen.

Partial rescue of defective spermatogenesis and infertility in RT-specific Sox17 cKO mice by AMH-Sox17 Tg

In RT-specific Sox17 conditional knockout (cKO) testes, deformation of the Sertoli valve (SV) leads to backflow of rete testis (RT) fluid into the seminiferous tubules, resulting in defective spermiogenesis and male infertility (Uchida et al., 2022). In contrast, AMH-Sox17 Tg mice exhibit excessive SV formation due to ectopic expression of SOX17 in SV Sertoli cells; however the AMH-Sox17 transgene is not activated in the RT epithelium. To evaluate whether this ectopic expression could rescue the SV defects in the RT-specific Sox17 cKO background, we examined the testis phenotype of double mutant mice (Sox17 cKO; AMH-Sox17 Tg, referred to as cKO;Tg) (Fig. 3). By crossing Tg; Sox17f/f males with SF1-Cre;Sox17f/f females, we generated four genotypes for comparison: control (Sox17f/f), cKO (SF1-Cre;Sox17f/f), Tg (Tg; Sox17f/f), and cKO;Tg (Sf1-Cre;Sox17f/f; Tg). As expected, control and Tg males showed normal spermatogenesis and the presence of spermatozoa in the epididymis. In contrast, cKO males exhibited abnormal spermiogenesis with a complete absence of epididymal spermatozoa (Fig. 3A), consistent with previous findings (Uchida et al., 2022). Interestingly, cKO;Tg double mutants showed improved spermatogenesis, with the presence of epididymal spermatozoa and a partial restoration of fertility compared to cKO males. Quantitative assessment of spermatogenesis using a 10-point scoring system (7 points for seminiferous tubule integrity + 3 points for presence of advanced germ cells) confirmed a significant improvement in cKO;Tg males (Fig. 3B) related to cKO males. This was further supported by an increased ratio of elongated spermatids in these cKO;Tg testes. Notably, epididymal spermatozoa were detected in 2 of 7 cKO;Tg individuals at 8 weeks of age. Furthermore, mating trials using four independent cKO;Tg males resulted in successful offspring production in 2 out of the 4 cKO;Tg males.

Morphometric analysis of spermatogenic quality in control (Sox17f/f), cKO (Sf1-Cre;Sox17f/f), Tg (Tg; Sox17f/f), and cKO;Tg (Sf1-Cre;Sox17f/f; Tg) testes

(A) H&E-stained sections of the epididymis and testes from control, cKO, Tg, and cKO;Tg males. Two representative testis images from both fertile and infertile cKO;Tg males are also presented. The lower panels present higher magnification images of damaged seminiferous epithelia. (B) Evaluation of spermatogenesis in cross-sections of convoluted seminiferous tubules (STs). The evaluation score (maximum 10 points) was calculated as the sum of the seminiferous tubule structural integrity score (Group A; 7 points) and the most advanced germ cell type score (Group B; 3 points). The two graphs on the right show the evaluation scores for each genotype (left; mean ± SD; *p<0.05; **p<0.01; one-way ANOVA) and the percentage of seminiferous tubules containing the most advanced germ cells (right). Defective spermatogenesis observed in cKO testes was rescued in the cKO;Tg double mutant testes. Scale bars: 50 μm (top two rows); 10 μm (bottom row) in A.

These results demonstrate that ectopic SOX17 expression in Sertoli cells partially rescues the structural and functional defects of the SV caused by the absence of SOX17 in the RT, thereby improving spermatogenesis and restoring fertility in otherwise infertile males.

Ectopic SOX17 expression in Sertoli cells rescues SV deformation in RT-specific Sox17 cKO testes

In RT-specific Sox17 conditional knockout (cKO) testes, disruption of the SV formation leads to defective spermiogenesis and male infertility (Uchida et al., 2022), which could be partially rescued by the ectopic SOX17 expression in Sertoli cells, as described above (Fig. 3). To assess whether such ectopic SOX17 expression could rescue the SV defects caused by Sox17 deficiency in the RT, we examined the RT-SV phenotypes of Sox17 cKO;AMH-Sox17 Tg double mutant (cKO;Tg) mice (Fig. 4). Double immunostaining for SOX17 and SOX9 confirmed that in cKO;Tg testes, SOX17 expression was restricted to Sertoli cells in the terminal segments of the seminiferous tubules, and was absent from the RT epithelium—opposite to the expression pattern seen in control testes, where SOX17 was confined to RT epithelial cells (Fig. 4A). In Tg testes, both RT epithelial cells and SV Sertoli cells were SOX17-positive. An expanded region of SOX17/SOX9 double-positive SV Sertoli cells was evident in Tg testes, while this region appeared to be reduced in cKO;Tg testes, suggesting partial normalization of the SV region (yellow area, Fig. 4A).

Partial normalization of the expanded SV structure in AMH-Sox17 Tg testes by cKO;Tg double mutants

(A) Immunostaining for SOX9 and SOX17 in the RT–SV–ST region of proximal testes in control, Tg, and cKO;Tg mice. Ectopic SOX17 expression is observed in Sertoli cells of Tg testes, whereas SOX17 expression is completely absent in RT cells of the cKO;Tg double mutants. The expanded region of SOX17/SOX9 double-positive SV Sertoli cells (yellow) seen in Tg testes appears to be reduced in cKO;Tg double mutants. (B) Immunostaining for SOX9 (green) and acetylated tubulin (SV marker; red) or E-cadherin (RT marker; red) demonstrates the restoration of proper SV formation in the RT–SV region of the cKO;Tg double mutants. Additional panels show immunostaining for SOX17 (red)/SOX9 (green) and HSP70 (elongated spermatid marker; red)/SCP3 (spermatocyte marker; green) in ST regions (left panels). (C) Quantitaive data of the nuclear density of Sertoli cells in the RT-SV region (mean ± SD; *p<0.05; **p<0.01; one-way ANOVA). RT, rete testis; SV, Sertoli valve. Scale bars: 200 μm (top row) and 50 μm (bottom row) in A and B.

Furthermore, acetylated tubulin-positive valve structures were properly formed in cKO;Tg testes, and normal spermatogenesis including SCP3-positive spermatocyte and Hsp70-positive elongated spermatids was observed in the convoluted seminiferous tubules (Fig. 4B). Quantitative analysis of SV Sertoli cell density revealed that the number of SOX9-positive nuclei in the RT-SV region was significantly increased in Sox17 Tg testes and significantly decreased in Sox17 cKO testes compared to control testes (Fig. 4C). Interestingly, in the cKO;Tg double mutant testes, SV Sertoli cell density was significantly higher than that in cKO testes, showing an intermediate level between Tg and cKO testes, comparable to that of controls (Fig. 4C). These findings indicate that ectopic SOX17 activity in SV Sertoli cells can rescue the structural defects of the SV in RT-specific Sox17 cKO testes, in a Sox17 dose-dependent manner.

Previous studies have shown that Rspo1 is highly expressed in RT epithelial cells in wild-type mice, but its expression is markedly reduced in cKO testes (Uchida et al., 2022). Along with the known expression of Wnt4 in SV Sertoli cells, we finally examined the expression patterns of Rspo1 and Wnt4 in the RT-SV region across different genotypes using RNAscope in situ hybridization (Fig. 5). Fixation, sectioning, and in situ hybridization were performed under identical conditions on serial sections. Expression profiles of RT-specific Rspo1 and SV-specific Wnt4 were generally comparable among control, Tg, cKO, and cKO;Tg mice (Fig. 5A), which confirm no ectopic Rspo1 expression in SOX17+ SV Sertoli cells. In Tg testes, Wnt4 expression appeared to be enhanced in the SV region, consistent with the expanded SV phenotype in these mice (Fig. 2). Notably, RT-specific Rspo1 expression was increased in Tg testes and slightly elevated in cKO;Tg mice relative to cKO mice, suggesting a potential paracrine mechanism, in which ectopic SOX17 activity in Sertoli cells might influence Rspo1 expression in adjacent RT epithelial cells.

Enhanced Rspo1 and Wnt4 expression in the RT-SV boundary of the testes in AMH-Sox17 Tg mice

RNA in situ hybridization showing the Rspo1 and Wnt4 expression profiles of RT-SV boundary in four genotypes: control (Sox17f/f), cKO (Sf1-Cre;Sox17f/f), Tg (Tg; Sox17f/f), and cKO;Tg (Sf1-Cre;Sox17f/f; Tg) testes. RT, rete testis; SV, Sertoli valve. Scale bar, 50μm.

Discussion

A previous study using RT-specific Sox17 cKO mice demonstrated that Sox17 expression is critical for Sertoli valve (SV) formation and proper spermiogenesis. Deletion of Sox17 in the rete testis (RT) disrupted the specification of modified Sertoli cells that form the SV, leading to backflow of RT fluid and abnormal detachment of immature spermatids within the seminiferous tubules (Uchida et al., 2022; Fig. 6). These findings established the essential role of SOX17+ RT epithelia in SV formation and clarified the physiological function of the SV in modulating fluid dynamics within the seminiferous tubules in vivo.

Schematic representation of valve structure size regulation by the AMH–Sox17 Tg mouse line

RT-specific Sox17 cKO mice exhibit deformed Sertoli valve (SV) structures, resulting in defective spermiogenesis and male infertility (Fig. 3; Uchida et al., 2022). Ectopic expression of SOX17 in Sertoli cells induces excessive SV formation in the terminal segments of the seminiferous tubule (ST), along with elevated Rspo1 expression in the rete testis (RT). Notably, this ectopic SOX17 expression partially rescues epididymal sperm production and restores fertility in cKO;Tg males. These Sox17 Tg mice, displaying SV hyperplasia and its functional rescue, provide valuable models for investigating paracrine signaling mechanisms within the RT-SV region.

In the present study, we demonstrated that ectopic expression of SOX17 in Sertoli cells induces excessive SV formation in the seminiferous tubules, along with increased Rspo1 expression in the RT. Moreover, this ectopic SOX17 expression partially rescued epididymal sperm production and restored fertility in cKO;Tg males (Fig. 6). These results suggest that SOX17 activity in the RT can be, at least in part, functionally substituted by its ectopic expression in Sertoli cells. The observed fertility recovery in cKO;Tg males, even in the absence of SOX17 in RT cells, implies that the primary role of SOX17 in the RT is restricted to SV formation in the terminal segments and is not essential for sperm transport and maturation in the RT-to-epididymis pathway in mice.

Ectopic SOX17 expression in Sertoli cells may mimic paracrine signaling normally mediated by SOX17+ RT epithelia in the SV region. One possible explanation is that RT epithelial cells and Sertoli cells share lineage similarities (Malolina and Kulibin, 2019; Uchida et al., 2022; Sato et al., 2026). Several key transcription factors of Sertoli cells —including WT1, NR5A1, SOX9, GATA4, and, in part, DMRT1— are also expressed in the RT epithelia (Aiyama et al., 2015; Malolina and Kulibin, 2019; Kulibin and Malolina, 2022; Uchida et al., 2022; Malolina et al., 2023). During sexual maturation, ectopic SOX17 expression in Sertoli cells may induce paracrine and/or autocrine signaling programs resembling those of RT cells, thereby promoting hyperplasia in Tg testes and facilitating SV recovery in cKO;Tg testes.

The Sox17 transgene is ectopically activated in Sertoli cells not only near the SV but also throughout the testis, whereas the SV phenotype is restricted to the RT-adjacent region of Tg testes (Figs. 1C, 2A, 4A). One possible explanation for this spatial discrepancy is that SOX17 may act on RT-lineage Sertoli-like cells located in the terminal segments adjacent to the RT. A recent study by Mayère et al. (2022) reported that PAX8+ RT progenitor cells, referred to as early supporting-like cells, contribute to the proximal region of the seminiferous tubules, suggesting heterogeneity in the origin of Sertoli cells near the RT during embryogenesis, even though PAX8 expression in fetal Sertoli cells is downregulated shortly after testis differentiation. It is therefore possible that ectopic SOX17 expression in these rare supporting-like cells may induce paracrine signaling similar to that of RT epithelia, thereby leading to SV phenotypes in Sertoli cells near the RT in Tg mice.

Interestingly, Rspo1 and Wnt4 signaling is activated in the RT-SV region (Uchida et al., 2022; this study). In Wnt4-null embryos, the RT region is underdeveloped at the seminiferous tubule–RT junction, leading to a failure in connecting the RT and seminiferous tubules, similar to the rete ovary (RO) phenotype (Mayère et al., 2022). These findings suggest that Wnt4–Rspo1 signaling is essential for proper connection of the terminal segment of seminiferous tubules to the RT and efferent ducts, which is critical for the integrity of the male reproductive tract. Our in situ hybridization analyses showed enhanced Rspo1 and Wnt4 expression driven by SOX17+ Sertoli cells in Tg mice (Fig. 4), which likely contributed to the rescue of SV formation observed in cKO;Tg mice (Fig. 5). The presence of protruding SV cells expressing Cyclin D1 supports the idea that SOX17-mediated paracrine activity may involve localized Rspo1-Wnt4 signaling during SV formation. Considering the potential involvement of other signaling pathways such as FGFs and TGFs, further studies are warranted to delineate the specific roles of individual paracrine factors in the RT–SV region prior to SV formation.

In conclusion, by generating a novel AMH-Sox17 Tg mouse line, we demonstrate that the size of the SV structure in the terminal segment of the seminiferous tubules can be experimentally manipulated. Furthermore, SV formation was successfully rescued in Sox17 cKO testes by introducing the AMH-Sox17 transgene into RT-specific Sox17 cKO mice. These genetically engineered models — exhibiting SV overdevelopment or restoration — provide valuable tools for investigating paracrine signaling and elucidating the biological significance of the SV and RT in the male reproductive system of mammals.

Materials and Methods

Animals

AMH-Sox17 Tg mice were generated by pronuclear injection of the plasmid-based transgene into the C57BL6 background, as described previously (Shinomura et al., 2014). The transgene was prepared by inserting the entire coding region of the Sox17 gene (Kanai-Azuma et al., 2002) downstream of the hAMH promoter-driven cassette (Fig. 1A). We obtained three independent F0 Tg males (lines #13, 26, and 27), among which a fertile mouse line #26 with a modest phenotype was mainly used in this study. F0 males of line #27 were infertile, possibly due to excessive detachment of the SV-like Sertoli cells with high SOX17 expression. The RT-specific Sox17 cKO mice (i.e., SF1-Cre;Sox17flox/flox mice; Dhillon et al., 2006; Kim et al., 2007) were produced from the Jackson Laboratory (Uchida et al., 2022). The SF1-Cre:Sox17flox/flox;AMH-Sox17 Tg double mutant (cKO;Tg), AMH-Sox17 Tg;Sox17 flox/flox (Tg), SF1-Cre;Sox17flox/flox (cKO) and Sox17 flox/flox (control) mice were mainly generated by crossing the AMH-Sox17 Tg; Sox17 flox/flox males with SF1-Cre:Sox17 flox/flox females. These mutant males at 4 and 8 weeks of age were used for the subsequent phenotypic analyses.

All animal experiments were performed following the Guidelines for Animal Use and Experimentation at the University of Tokyo, and all experimental procedures performed herein were approved by the Institutional Animal Care and Use Committee, Graduate School of Agricultural and Life Sciences, University of Tokyo (approval IDs: P13-762, P13-764, P22-147). The PCR list for the genotyping was shown in Table S1.

Histology and immunohistochemistry

Testes were fixed in 4% PFA at 4°C overnight, dehydrated through an ethanol series, and embedded in paraffin. The paraffin sections (4 μm thickness) were subjected to hematoxylin and eosin (H&E) staining, PNA lectin (acrosome) staining and immunohistochemistry. For immunohistochemistry, the sections were incubated overnight at 4°C with the primary antibodies listed in Table S2. Immunoreactions were visualized using Alexa fluor-488/555/594/680-conjugated secondary antibodies (Abcam). Samples were analyzed by fluorescence microscopy (BX51N-34-FL-2) or Leica TCS SP8 confocal laser microscopy. Testes from at least four animals in each group were examined, and tissue sections from cKO, Tg and cKO;Tg and their littermate controls were processed in parallel to ensure the reproducibility of the results.

Electron microscopy

Testes were trimmed and fixed in 2.5% glutaraldehyde in 0.1 M sodium cacodylate buffer (pH 7.4) overnight, post-fixed with 0.5% OsO4 suspended in 0.1 M phosphate buffer (pH 7.3) for 30 min, and dehydrated through a graded ethanol series. After passing through propylene oxide, the tissues were embedded in Epon 812. Ultrathin sections were cut, stained with uranyl acetate and lead citrate, and observed by TEM (JEM-1010; JEOL).

RNA in situ hybridization

Testes were fixed in 10% formalin, dehydrated through an ethanol series, and embedded in paraffin. The paraffin sections (4 μm thickness) were processed for RNA in situ detection using the RNAscope 2.5HDAssay Red Manual (ACDBio) as described previously (Imura-Kishi et al., 2021; Uchida et al., 2022). The probes used in this study are listed in Table S3.

Evaluation of Spermatogenesis score

To evaluate spermatogenesis in mouse testes, we performed the fluorescence staining with PNA lectin (a marker for acrosome; Nakata et al., 2015) on tissue sections (Control: n = 4; cKO: n = 4; Tg: n = 5; cKO;Tg: n = 5 males). Fluorescent images of entire testis sections were acquired using a confocal microscope (TCS SP8 Leica Microsystems) with the Tiling function in LAS X software. A two-axis scoring system was applied: Group A categorized the morphological condition of the seminiferous tubules, and Group B assessed the most advanced germ cell stage present within each tubule. Scores were assigned as maximun 10 points as follows: Group A evaluated the morphological condition of the seminiferous tubules, while Group B assessed the most advanced germ cell stage present in each tubule. Scores were assigned with a maximum of 10 points (Group A [tubule condition]: normal [7 points]; disrupted spermatocyte stage [5 points]; partial cell dropout or loss [3 points]; Sertoli cells only [0 points]) plus Group B [most advanced germ cell type]: elongated spermatids [3 points]; round spermatids [2 points]; spermatocytes [1 point]; spermatogonia only [0 points]) (see Fig. 3B).

Quantitative morphometric analyses

The number of Sertoli cells in the SV region was analyzed by using SOX9/E-cad/DAPI immunostained sections. In brief, the nuclear density of the SOX9-positive/E-cad-negative cells within RT-SV region was counted in each terminal segment of the seminiferous tubule (~100 μm from the edge of the RT; Fig. 4C). Only the SV sectioned at the mid-sagittal plane was subjected to testicular section analysis.

Statistical analysis

For comparisons among four genotype groups, one-way ANOVA followed by Tukey’s multiple comparisons test was performed. Quantitative data are presented as means ± standard deviation (SD) and shown as bar graphs with dot plots. Statistical analysis was performed using GraphPad Prism (v.10.4.2; GraphPad Software, San Diego, CA, USA). P < 0.05 was considered indicative of significance, and levels of significance are shown as *P < 0.05 and **P < 0.01. In addition, single comparisons between two groups, as shown in Fig. S1A, were were analyzed using a two-tailed Student’s t-test (data are presented as means ± standard error of the mean).

eLife Assessment

This important study clearly demonstrates that Sox17 is key for the formation and function of the Sertoli valve, a transition region between the rete testis and seminiferous tubules, which remains an understudied domain of testicular biology. The supporting data are generally convincing but remain incomplete. This work will be of interest to reproductive biologists and andrologists who work on male fertility and men's health.

Reviewer #1 (Public review):

Summary:

This manuscript is an excellent follow-up to your 2022 study, in which Sox17 expression was localized to the rete testis and shown to be required for proper formation of the Sertoli cell valve (transition region). By using Nr5a1-Cre to drive conditional deletion of Sox17 specifically in rete testis cells, you demonstrate that testis weights remain normal at 2 weeks of age but become significantly reduced by 8 weeks in Sox17-cKO males. At the later time point, the seminiferous epithelium is severely disrupted, with apparent arrest of spermiogenesis: the epididymal lumen is essentially devoid of sperm, and most tubules lack elongated spermatids.

Strengths:

The study clearly shows the role of Sox17 in Sertoli cells as being important to SV function. The SV (transition region) between the rete testis and seminiferous tubules remains an understudied domain of testicular biology. The present work, together with the authors' prior study, highlights intriguing mechanisms operating in this specialized niche.

Weaknesses:

At the same time, the available data do not yet fully explain either the developmental assembly of the Sertoli valve or the precise consequences of its functional disruption. These studies are nonetheless valuable precisely because they raise more questions than they answer; the conceptual implications are thought-provoking.

Reviewer #2 (Public review):

This manuscript investigates the role of SOX17 in the formation and function of the Sertoli valve (SV) at the interface between seminiferous tubules and the rete testis (RT). Building on previous work showing that rete testis-specific deletion of Sox17 disrupts SV formation, leading to defective spermiogenesis and male infertility, the authors explore how SOX17 overexpression in Sertoli cells regulates the SV of rodent testes.

Using transgenic mouse models with ectopic Sox17 expression in Sertoli cells, the study demonstrates that SOX17 is not only required but can also modulate SV formation. Ectopic expression in Sertoli cells induces expansion of the SV structure and partially rescues SV defects and spermatogenesis in RT-specific Sox17 conditional knockout animals. The data support a model in which SOX17 acts through paracrine signaling to regulate SV formation, although the precise mechanisms remain to be clarified.

Overall, this is a well-executed study with novel and significant findings. The ability to experimentally manipulate SV size is particularly compelling and provides a valuable framework to study fluid dynamics and epithelial interactions in the testis. This work will be of broad interest to the reproductive biology and developmental biology communities.

Reviewer #3 (Public review):

Summary:

These studies are based on previously published work that showed that deletion of expression of the Sox17 gene in the testis essentially deleted the formation of the Sertoli valve in the Rete testis. The authors extended this work by constructing a vector that resulted in increased Sox17 expression by Sertoli cells and enhanced formation of the Sertoli valve in both wild type and Sox17 knockout mice. The work provides strong evidence supporting the requirement for Sox17 expression to allow formation of the Sertoli valve.

Strengths: The general approach was to express Sox17 from a Tg mouse that expressed Sox17 from Sertoli cells. This Tg mouse was bred into both the WT and the Sox17 KO mouse. The Sertoli valve was enhanced in both the WT/Tg mouse and KO/Tg mouse, showing that ectopic Sox17 could compensate in the Sox17 Ko and act in a concentration-dependent manner in the WT mouse. The results are strong and support the conclusions from the authors. The results were as expected from the original paper describing the KO of Sox 17. These results strengthen these conclusions and provide ideas for additional conclusions. These studies were technically challenging, and the authors provided a very solid manuscript.

Weaknesses:

The authors refer several times to high or low expression, but it all appears to be based on immunohistochemistry, and there is no real quantification using PCR, for example. The process used for cell quantification lacks a rationale for why certain numbers were assigned.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public review):

Summary:

This manuscript is an excellent follow-up to your 2022 study, in which Sox17 expression was localized to the rete testis and shown to be required for proper formation of the Sertoli cell valve (transition region). By using Nr5a1-Cre to drive conditional deletion of Sox17 specifically in rete testis cells, you demonstrate that testis weights remain normal at 2 weeks of age but become significantly reduced by 8 weeks in Sox17-cKO males. At the later time point, the seminiferous epithelium is severely disrupted, with apparent arrest of spermiogenesis: the epididymal lumen is essentially devoid of sperm, and most tubules lack elongated spermatids.

Strengths:

The study clearly shows the role of Sox17 in Sertoli cells as being important to SV function. The SV (transition region) between the rete testis and seminiferous tubules remains an understudied domain of testicular biology. The present work, together with the authors' prior study, highlights intriguing mechanisms operating in this specialized niche.

Weaknesses:

At the same time, the available data do not yet fully explain either the developmental assembly of the Sertoli valve or the precise consequences of its functional disruption. These studies are nonetheless valuable precisely because they raise more questions than they answer; the conceptual implications are thought-provoking.

Reviewer #2 (Public review):

This manuscript investigates the role of SOX17 in the formation and function of the Sertoli valve (SV) at the interface between seminiferous tubules and the rete testis (RT). Building on previous work showing that rete testis-specific deletion of Sox17 disrupts SV formation, leading to defective spermiogenesis and male infertility, the authors explore how SOX17 overexpression in Sertoli cells regulates the SV of rodent testes.

Using transgenic mouse models with ectopic Sox17 expression in Sertoli cells, the study demonstrates that SOX17 is not only required but can also modulate SV formation. Ectopic expression in Sertoli cells induces expansion of the SV structure and partially rescues SV defects and spermatogenesis in RT-specific Sox17 conditional knockout animals. The data support a model in which SOX17 acts through paracrine signaling to regulate SV formation, although the precise mechanisms remain to be clarified.

Overall, this is a well-executed study with novel and significant findings. The ability to experimentally manipulate SV size is particularly compelling and provides a valuable framework to study fluid dynamics and epithelial interactions in the testis. This work will be of broad interest to the reproductive biology and developmental biology communities.

Reviewer #3 (Public review):

Summary:

These studies are based on previously published work that showed that deletion of expression of the Sox17 gene in the testis essentially deleted the formation of the Sertoli valve in the Rete testis. The authors extended this work by constructing a vector that resulted in increased Sox17 expression by Sertoli cells and enhanced formation of the Sertoli valve in both wild type and Sox17 knockout mice. The work provides strong evidence supporting the requirement for Sox17 expression to allow formation of the Sertoli valve.

Strengths:

The general approach was to express Sox17 from a Tg mouse that expressed Sox17 from Sertoli cells. This Tg mouse was bred into both the WT and the Sox17 KO mouse. The Sertoli valve was enhanced in both the WT/Tg mouse and KO/Tg mouse, showing that ectopic Sox17 could compensate in the Sox17 Ko and act in a concentration-dependent manner in the WT mouse. The results are strong and support the conclusions from the authors. The results were as expected from the original paper describing the KO of Sox 17. These results strengthen these conclusions and provide ideas for additional conclusions. These studies were technically challenging, and the authors provided a very solid manuscript.

Weaknesses:

The authors refer several times to high or low expression, but it all appears to be based on immunohistochemistry, and there is no real quantification using PCR, for example. The process used for cell quantification lacks a rationale for why certain numbers were assigned.

We sincerely thank the reviewers for their careful evaluation of our manuscript and for their constructive and encouraging comments. We are grateful for the recognition of the significance of the Sertoli valve as an understudied transition region between the rete testis and seminiferous tubules, as well as for the positive assessment of our genetic approach and the evidence that ectopic SOX17 expression can modulate SV formation. We have carefully considered all points raised in the assessment and have revised the manuscript accordingly. The major revisions include:

(1) Clarification of the scope and limitations of the study (Reviewers #1 and #2):

In response to the comments that the developmental assembly of the Sertoli valve and the precise consequences of its functional disruption remain incompletely understood, we clarified the scope and limitations of the present study at the end of 7th paragraph in the Discussion. Although our findings support a model in which SOX17 regulates SV formation through paracrine signaling, the downstream effectors and precise molecular mechanisms remain to be identified. We therefore revised the Discussion to avoid overinterpretation of the molecular mechanisms and to emphasize that comprehensive mechanistic analyses, including transcriptomic analyses using the Tg mouse model, represent an important direction for future research. We also added histological analyses of the earliest detectable lesions at 4 weeks of age and low-magnification images of adult Sox17 cKO testes (new Figure S1), revealing selective sloughing of round spermatids despite preserved Sertoli cell architecture and subsequent mosaic spermatogenic defects among individual seminiferous tubules. These observations provide additional insights into the altered luminal microenvironment and suggest that spermatogenic defects may progress in a tubule-by-tubule manner.

(2) Clarification of quantitative analysis and methodology (Reviewer #3):

In response to concerns regarding the basis and methodology of cell quantification, we revised the Methods to provide detailed information on tissue preparation, fixation, orientation of the rete testis–Sertoli valve region, and the criteria used for quantitative analysis of SV-associated Sertoli cells (new Figure S4). We clarified that Sertoli cells were counted within the SV region extending approximately 100 μm from the RT boundary, including Sertoli cells protruding into the RT lumen, based on previously established criteria (Aiyama et al., 2015).

(3) Clarification of the limitations of expression-level assessment (Reviewer #3):

In response to concerns regarding the quantitative assessment of SOX17 and other SV-associated molecules, we clarified the technical limitations of selectively isolating the very small SV region and obtaining sufficient material for quantitative molecular analyses such as qPCR at the end of 7th paragraph in the Discussion. We therefore clarified that expression of SV-associated molecules in the present study was primarily evaluated using histological and immunohistochemical approaches and added the relevant text to acknowledge these limitations.

We also made additional revisions to clarify each mouse Tg line, phenotypic descriptions, standardize gene nomenclature, improve methodological descriptions, and refine the relevant Discussion where appropriate.

We sincerely appreciate the reviewers’ thoughtful and constructive comments. Their feedback has helped us clarify the scope of our conclusions, strengthen the methodological descriptions, and improve the overall presentation of the study. All changes have been incorporated into the revised manuscript.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

(i) Although the current paper is not responsible for interpreting the 2022 findings, both datasets show reduced spermatid production accompanied by multinucleated giant germ-cell syncytia. This phenotype has been attributed to backflow of tubular fluid and consequent microenvironmental perturbation. While this is a reasonable hypothesis, it is not entirely consistent with earlier experimental observations. Complete ligation of the efferent ductules reliably produces giant cells, whereas estrogen-receptor knockout, which also causes massive luminal fluid accumulation, does not. In addition, ligation of the testicular artery itself can induce giant-cell formation. Although this may have already been answered in the papers, can you be sure that a direct or indirect effect on the vasculature can be excluded in the Sox17-cKO model?

We thank the reviewer for this important comment. In our models, SOX17 expression was manipulated specifically in the Sertoli cell lineage, either by SF1-Cre-mediated Sox17 deletion or by ectopic SOX17 expression under the hAMH-promoter. SOX17-expressing vascular endothelial cells were not targeted in either model, making a direct effect of Sox17 manipulation on the testicular vasculature unlikely. Moreover, the partial rescue of the Sox17 cKO phenotype by hAMH-Sox17 supports the interpretation that the phenotype primarily results from altered SOX17 function in Sertoli cells and RT epithelia.

However, indirect effects on the vascular or interstitial environment by aberrant luminal flow cannot be completely excluded, particularly with the substantial accumulation of sloughed round spermatids (giant cells) within the rete testis. Addressing the potential for an initial luminal flow defect, we newly added histological images of 4-week-old testes (Figure S1), where selective post-meiotic germ cell sloughing occurs despite preserved Sertoli cell process architecture, suggesting an altered adluminal microenvironment that impairs Sertoli–spermatid adhesion. Furthermore, low-magnification images of adult mature Sox17 cKO testes (Figure S1B) display a mosaic pattern of spermatogenic defects across individual tubules. While 3D reconstruction was not conducted, this structural pattern supports the view that spermatogenic failure progresses on a tubule-by-tubule basis, potentially linked to the structural integrity of individual Sertoli valves.

(ii) A related and important unresolved issue is the total number of Sertoli cells per testis in cKO males. The number of Sertoli cells per tubule cross-section is reported to be equivalent to controls; however, the substantial reduction in testis weight implies a corresponding reduction in tubule length. Under these conditions, maintenance of a normal per-cross-section count would still be compatible with an overall decrease in total Sertoli-cell number. Although it is generally accepted that murine Sertoli cells exit the cell cycle around postnatal day 15, continued growth of the testis may still occur in the Sertoli valve region, where Sertoli cells retain proliferative capacity. Your discussion of possible heterogeneity in the embryonic origin of Sertoli cells near the rete testis is therefore particularly intriguing and commendable. Should this hypothesis be substantiated, it would raise the possibility that Sertoli cells derived from the valve region, especially those that migrate into the seminiferous tubules, are intrinsically less competent to support full spermatogenesis than those of classic gonadal-ridge origin.

To help readers appreciate the overall severity and topographic distribution of the spermatogenic defect (particularly in tubule segments distant from the rete), inclusion of a low-magnification photomicrograph of a well-fixed (Bouin's) testicular cross-section would be very useful.

We thank the reviewer for this important comment. We agree that the maintenance of Sertoli cell numbers per seminiferous tubule cross-section does not necessarily indicate preservation of the total Sertoli cell number per testis, particularly given the substantial reduction in testis size and potential reduction in overall seminiferous tubule length. Although total Sertoli cell numbers can theoretically be estimated using stereological approaches, such analyses are technically demanding and beyond the scope of the present study.

We also appreciate the reviewer’s insightful suggestion regarding potential heterogeneity among Sertoli cell populations. Sertoli cells associated with the Sertoli valve region may have distinct developmental origins or functional properties compared with classical gonadal ridge-derived Sertoli cells, which could potentially influence their capacity to support complete spermatogenesis. Although this hypothesis was not directly tested in this study, we have expanded the Discussion to highlight the developmental and functional heterogeneity of Sertoli cell populations associated with the Sertoli valve as an important topic for future investigation.

In addition, as requested, we have added a low-magnification image of well-preserved testicular cross-sections in Supplementary Figure S1B to better illustrate the overall severity and topographic distribution of spermatogenic defects throughout the testis.

Specific Comments:

(1) Figure 3A and associated fertility/histology data. The results state that epididymal spermatozoa were detected in only 2 of 7 cKO;Tg males at 8 weeks of age, yet Materials and Methods indicate that spermatogenesis was evaluated in only 5 males. a) Were the remaining two males also examined histologically? b) It would be interesting to determine if the severity of pathological changes was the same in regions more distant from the rete testis, or possibly different tubules. See: Nakata H, Wakayama T, Sonomura T, Honma S, Hatta T and Iseki S (2015). "Three-dimensional structure of seminiferous tubules in the adult mouse." J Anat 227(5): 686-694. c) In addition, mating trials were performed with four independent cKO;Tg males, two of which sired offspring. It is unclear whether the testes of these four mating males were included among the five (or seven) animals evaluated for histology, and whether the two fertile males correspond exactly to the two individuals that retained epididymal sperm. Please clarify these relationships explicitly so that readers can correctly interpret the link between histological findings and fertility.

We thank the reviewer for this important comment. We apologize that the relationship among the groups of animals used for histological analysis, epididymal sperm detection, and fertility assessment was not sufficiently clear in the original manuscript. Because this study focused specifically on the anatomically minute RT–SV region, our sampling strategy had to prioritize the maximal utilization of this limited tissue. In this study, the RT–SV region, the remaining testicular tissue, and the epididymis were processed separately as three tissue blocks for each animal (Figure S4) and were independently evaluated for distinct analysis sets. Briefly, the proximal quarter containing the rete testis and Sertoli valve region was used for SV analysis, whereas the remaining three-quarters of the testis were used for evaluation of spermatogenesis, and the epididymis was analyzed separately for the presence of spermatozoa. Therefore, due to these technical requirements, tissue allocation, and independent analytical evaluation, the numbers of animals used for RT–SV analysis, testicular histology, epididymal sperm detection, and fertility testing were not identical.

For quantitative histological analyses, we also used virgin males to minimize potential variation associated with mating experience and to allow comparison with age-matched littermate controls. Therefore, these animals were not used for fertility testing. Fertility assessment was performed using an independent cohort of cKO; Tg males that were subjected to long-term mating trials with wild-type females. Thus, fertility outcomes and histological findings were not designed to be directly matched at the individual level.

In response to the reviewer’s suggestion, we have clarified the selection of experimental animals and the relationship among fertility assessment and histological analyses in the Materials and Methods and added a schematic illustration of the sampling strategy in Figure S4. We also corrected the citation for Nakata H et al., 2015 in the revised manuscript.

(2) Page 8, line 301 (Sertoli-cell quantification). The description of the counting method-"counted in each ... (~100 μm from the edge of the RT; Fig. 4C)"-is ambiguous.

(a) Does this mean that cells were counted beginning at the rete boundary and extending radially outward for approximately 100 μm, or is a circumferential sampling area intended? (b) Figure 4C shows a large standard deviation, indicating substantial variability with the current approach. An alternative strategy (for example, counting Sertoli cells within standardized areas or per tubule specifically within the valve region) might reduce variability and improve reproducibility. Regardless of the method ultimately chosen, a more precise, step-by-step description of the quantification protocol is required so that it can be reliably replicated by other laboratories.

We thank the reviewer for pointing out that the description of the Sertoli cell quantification method was not sufficiently clear. The Sertoli cell quantification was performed using the same criteria as previously described (Aiyama et al., 2015; Uchida et al., 2022), in which SOX9-positive Sertoli cell nuclei within the SV-associated region were counted.

In the revised manuscript, we have clarified that the SV region was operationally defined as comprising (i) the terminal 100 μm segment of the seminiferous tubule immediately adjacent to the rete testis (RT) and (ii) the protruded SV extending into the RT lumen. Based on the distribution of spermatogonial stem cells, the ~100 μm region extending from the RT boundary along the seminiferous tubule toward the ST side was defined as the SV region (Aiyama et al., 2015). Only sagittal sections showing a continuous RT–SV–ST axis and sectioning the SV approximately through its mid-sagittal plane were included for quantitative analysis.

Furthermore, to improve reproducibility, we have added a more detailed description of the tissue preparation and quantification procedures in the Materials and Methods and provided a schematic illustration of the quantification strategy in the new Figure S4.

Reviewer #2 (Recommendations for the authors):

(1) Phenotypic differences between transgenic lines: the phenotypic differences between the tg26 and tg27 lines are intriguing and warrant further clarification. While tg27 mice exhibit infertility and defective spermatogenesis, tg26 animals remain fertile with SV expansion. Could the authors elaborate on the underlying causes of these differences? In particular, is infertility in tg27 mice due to excessive SOX17 expression impairing Sertoli cell function? A comparison of Sox17 expression levels between tg26 and tg27 lines would be informative. In addition, it would be useful to assess whether acetylated tubulin (Ac-Tub) expression is present in the Sertoli cells of the tg27 mouse testis.

We thank the reviewer for this highly constructive and insightful comment. We clarified in the revised manuscript that the analysis of the Tg27 mouse was performed using the F0 founder male and added an explanation that only the Tg26 line could be established as its heterogenous SOX17 expression in Sertoli cells did not impair overall fertility. We agree that the phenotypic differences between the Tg26 line and the Tg27 mouse provide important clues regarding the dosage-dependent effects of SOX17 in Sertoli cells. Unfortunately, we were unable to establish a stable, multi-generational transgenic line from this Tg27 founder (F0) male. Consequently, we could not perform detailed molecular or immunohistochemical analyses on this line beyond the initial histological evaluation of the F0 generation presented in Figure 1. For this reason, we cannot provide a quantitative comparison of Sox17 expression levels or evaluate acetylated tubulin (Ac-Tub) expression in Tg27 Sertoli cells.

To address the reviewer's concern without overstepping the available data, we removed direct quantitative comparisons of Sox17 expression levels between the two lines from the text. Instead, we added a clear description of their contrasting cellular expression patterns - specifically, the mosaic, heterogeneous SOX17 expression in Tg26 Sertoli cells versus the ectopic, uniform SOX17 expression in the infertile #27 F0 male - in the 'Animals' section of Materials and Methods. This mosaic pattern in Tg26 testes suggests the presence of Sertoli cells with low or undetectable SOX17 levels, which may be associated with sustaining overall fertility.

(2) Mechanism of SOX17 action: although SOX17 is a transcription factor, the author's studies indicate it regulates SV formation via paracrine and/or autocrine signaling. The underlying mechanisms remain unclear. Which downstream factors mediate this effect? The observed upregulation of RSPO1 and WNT4 is suggestive, but more direct evidence would strengthen this conclusion. For example, does SV expansion in tg26 mice depend on the activation of RSPO1/WNT signaling? Additional molecular analyses, such as bulk RNA-seq comparing control and transgenic testes, could help identify pathways regulated by SOX17 and clarify its mode of action.

We thank the reviewer for this important and insightful suggestion. At present, comprehensive analyses, including scRNA-seq of Sox17 cKO and littermate control testes, have not identified definitive downstream targets of SOX17 (Uchida et al., 2022). As the reviewer rightly points out, the Tg26 mouse model generated in this study represents a valuable tool for investigating SOX17-dependent molecular pathways. To this end, we are currently conducting transcriptomic analyses of Tg26 seminiferous tubules to identify genes altered in SOX17+ Sertoli cells. However, determining whether these candidate genes represent direct transcriptional targets of SOX17 and whether they function specifically in the rete testis-associated region during Sertoli valve formation will require extensive functional and expression studies. Therefore, we feel it would be premature to draw definitive conclusions regarding the underlying molecular mechanisms, including the precise involvement of the RSPO1/WNT signaling pathway, in the present manuscript. Accordingly, rather than overinterpreting the available data, we have revised the Discussion to clarify this limitation (at the end of 7th paragraph in the Discussion). Furthermore, incorporating initial insights from our ongoing Tg26 transcriptomic analyses, we have added a brief discussion, supported by relevant literature, on the possibility that SOX17 may regulate Sertoli valve formation by modulating cell adhesion and extracellular matrix (ECM) organization and altering the responsiveness of SOX17-positive Sertoli cells to morphogenetic signals originating from the rete testis (new 5th paragraph in Discussion).

(3) Minor comment: Gene nomenclature should be standardized: e.g. line 245, Sox17 and hAMH should be italicized.

We thank the reviewer for pointing this out. All gene names have been italicized throughout the manuscript.

Reviewer #3 (Recommendations for the authors):

No suggestions except to quantify some of the changes in concentration of agents by PCR rather than eyeball levels with immunocytochemistry. Verify the cell quantification procedure used.

We thank the reviewer for this comment. The Sertoli valve (SV) is an extremely small transitional structure, with only approximately 20 sites per mouse testis. As a result, selective isolation of the SV region to collect sufficient material for molecular analyses, such as quantitative PCR, remains technically challenging. We have therefore added this limitation to the Discussion.

Regarding the cell quantification procedure, we have clarified the methodology in the revised Materials and Methods and added a schematic illustration in Figure S4. Specifically, the Sertoli cell number in the SV region was quantified by counting SOX9-positive Sertoli cell nuclei within a standardized SV-associated region in RT–SV–ST sagittal sections.

Version 2

Abstract

In the terminal segment of the seminiferous tubules, SOX17 expression in the rete testis (RT) epithelium plays a crucial role in the formation of the Sertoli valve (SV), as revealed by phenotypic analyses of RT-specific Sox17 conditional knockout (cKO) mouse testes. In these RT-specific Sox17 cKO testes, SV disruption leads to the backflow of RT fluid into the seminiferous tubules, resulting in defective spermiogenesis and male infertility. Although valve deformation in the Sox17 cKO testes is likely caused indirectly by impaired downstream actions of Sox17 in the RT, the mechanisms by which SOX17 in RT influences SV formation in the seminiferous tubules remain unclear. To address this, we generated a novel AMH-Sox17 transgenic (Tg) mouse line carrying a human AMH promoter-driven Sox17 cDNA cassette. We analyzed the phenotypes of the Sertoli valve and spermatogenesis in AMH-Sox17 Tg mice, as well as in RT-specific Sox17 cKO; AMH-Sox17 Tg double mutant mice. Ectopic SOX17 (SOX17+) expression in Sertoli cells resulted in excessive Sertoli valve structures with acetylated tubulin bundles in the terminal segment of the AMH-Sox17 Tg testes, along with enhanced WNT4/RSPO1 signaling, suggesting the enhanced valve formation of ectopic SOX17+ Sertoli cells by themselves. Moreover, the AMH-Sox17 Tg could partially rescue the SV deformation and infertility in RT-specific Sox17 cKO mice, leading to proper SV formation, normal spermiogenesis and a partial recovery of male fertility in AMH-Sox17 Tg; RT-specific Sox17 cKO double mutant mice. These findings genetically demonstrate that ectopic SOX17+ Sertoli cells can compensate for SOX17 paracrine signaling in the RT, underscoring a key shared downstream pathway between RT and SV.

Introduction

Valves are critical anatomical structures that ensure unidirectional fluid flow in the circulatory and lymphatic systems. They function passively to prevent backflow, maintaining efficient fluid dynamics within vessels (Moore and Bertram, 2018; O’Donnell and Yutzey, 2020). When valve integrity is compromised, it distorts the luminal architecture, reduces intravascular pressure, and may result in regurgitation.

In the mammalian testis — an organ characterized by active, oscillatory fluid flow (Fleck et al., 2021)— a valve-like structure called the Sertoli valve (SV) emerges at the termini of seminiferous tubules (STs) during puberty (Aiyama et al., 2015; Nagasawa et al., 2018; Imura-Kishi et al., 2021). The SV comprises a specialized subset of Sertoli cells that exhibit elevated levels of phosphorylated AKT and STAT3. These cells extend cytoplasmic projections enriched with acetylated tubulin-positive (ace-TUB⁺) microtubules into the rete testis (RT) lumen, forming a dynamic and structurally distinct valve-like barrier (Russell, 1993). Recently, we identified SOX17, a transcription factor expressed in RT epithelial cells, as essential for SV formation and for spermiogenesis in the ST. In mice with RT-specific conditional knockout of Sox17 (Sox17 cKO), the SV fails to form properly before puberty, allowing retrograde fluid flow from the RT into the convoluted ST (Uchida et al., 2022). Dysfunction of the RT–SV valve disrupts spermatogenesis in the upstream seminiferous tubules, leading initially to spermatid sloughing with preservation of the Sertoli cell architecture, followed by the progressive involvement of individual seminiferous tubules with age (Uchida et al., 2022; see Fig. S1). These findings imply the critical role of the RT-SV structure in maintaining proper fluid dynamics essential for spermatogenesis.

Interestingly, the SV epithelium provides a specialized niche for undifferentiated spermatogonia (Aundiff) and immature/proliferative Sertoli cells (Aiyama et al., 2015; see reviews by Figueiredo et al., 2021; Seco-Rovira et al., 2025). The basal compartment of the SV epithelium supports self-renewal of Aundiff spermatogonia by maintaining high expression of growth factors such as FGFs (Takashima et al., 2015; Kitadate et al., 2019; Yang et al., 2021; also see the review by Hofmann and McBeath, 2022) and GDNF (Meng et al., 2000; Uchida et al., 2016), together with by locally suppressing retinoic acid (RA)-induced differentiation (Imura-Kishi et al., 2021). Furthermore, SV Sertoli cells display Cyclin D1-positive states in mice and hamsters, and in rats and hamsters, some of these cells proliferate and contribute to the convoluted ST even at adult stage (Aiyama et al., 2015; Figueiredo et al., 2019), suggesting that the SV acts as a niche for proliferative Sertoli cells in hamsters and rats. Specification of these SV Sertoli cells is thought to be regulated by paracrine signals such as RSPO1, FGFs etc. from adjacent SOX17-positive RT cells (Imura-Kishi et al., 2021; Uchida et al., 2022; see review by Han et al., 2026). However, how SOX17-positive RT cells regulate SV formation in the proximal region of rodent testes remains unclear.

To address this question, we generated a novel transgenic (Tg) mouse line that exhibits excess SV formation by the Sertoli cells ectopically expressing Sox17 under the control of the human AMH promoter (SOX17+ Sertoli cells). Such SOX17+ Sertoli cells enhance the expression of WNT4 and RSPO1 in SV and RT regions, respectively. Moreover, we showed a partial rescue by the SOX17+ Sertoli cells in the Sertoli valve formation and aberrant spermiogenesis in the cKO testes with Sox17-null RT. These findings highlight the importance of SOX17-downstream paracrine actions in the SV and RT for the proper valve formation, rather than the SOX17 roles in the RT formation in itself.

Results

Overexpression of SOX17 in Sertoli cells induced hyperplasia of the SV region

To investigate the downstream paracrine effects of Sox17, we generated AMH-Sox17 transgenic (Tg) mice that ectopically express Sox17 in Sertoli cells under the control of the human anti-Müllerian hormone (AMH) promoter (Fig. 1A, S2). Among three Tg founders, one F0 Tg male (#27) exhibited infertility and was smaller in size of testes compared to his wild-type littermates (Fig. 1B). Histologically, the most prominent abnormalities were confined to the terminal region of the seminiferous tubules, whereas the overall architecture and spermatogenesis of the remaining seminiferous tubules were largely preserved (Fig. 1C). In this Tg male, the Sertoli valve (SV) structure appeared to be expanded in the terminal segments of the seminiferous tubules (Fig. 1B, C). Additionally, abnormal aggregates of Sertoli cells were frequently observed on the luminal side of some seminiferous tubules near the RT (indicated by asterisks in Fig. 1B). Immunostaining for SOX17 and E-cadherin in the #27 Tg testis revealed ectopic and uniform expression of SOX17 in Sertoli cells throughout both the convoluted seminiferous tubules and the SV-like regions, similar to those observed in the RT epithelium (Fig. 1D). Extensive detachment of SV-like Sertoli cells was also observed near the RT, and the #27 F0 Tg male was infertile.

Hyperplastic SV structure in the AMH-Sox17 Tg male mouse (F0)

(A) Schematic diagram of the DNA construct containing Sox17 coding sequences driven by the human AMH promoter. (B, C) Gross morphology and hematoxylin and eosin (H&E)-stained sections of testes from infertile Tg #27 and its wildtype littermates. In the Tg testis, normal spermatogenesis is partially observed in the convoluted seminiferous tubules (ST), while an expanded SV-like region including decidual cell aggregates is evident in the proximal area near the rete testis (RT) (asterisks in C). (D) Immunostaining for SOX17 and E-cadherin (E-cad; a marker for both RT epithelia and Aundiffer spermatogonia), showing ST regions (left two panels) and SV regions (rightmost panel) of the testis sections. The AMH-Sox17 Tg testis exhibits ectopic expression of SOX17 in almost all of the Sertoli cells in both ST and SV. Asterisks in (D) indicate protruded SOX17-positive SV Sertoli cells within the E-cad-positive RT lumen. RT, rete testis; SV, Sertoli valve. Scale bars: 1 mm (B), 200 μm (C), 50 μm (D).

Establishment of the AMH-Sox17 Tg Line (#26) with Excess SV Formation

One Tg male (#26) was fertile, and the AMH-Sox17 transgene was successfully transmitted to its offspring (Fig. S2A, B). The morphology and weight of the testes in Tg mice were comparable to those of their wild-type littermates. Although the AMH-Sox17 transgene was activated in a subset of Sertoli cells during both fetal and postnatal stages (Fig. S2C, D), an expansion of the acetylated tubulin-positive Sertoli valve (SV) region was observed in the proximal testis of this line (Fig. 2A), similar to that seen in the infertile #27 F0 Tg male with a more severe phenotype (Fig. 1D). Because this Tg line could be stably maintained by backcrossing beyond the 8th generation, line #26 was used for all subsequent experiments (hereafter referred to as “Tg”).

Morphological characterization of the hyperplastic SV structure in a novel AMH-Sox17 Tg line

(A) Immunofluorescence staining for acetylated tubulin (red) and cyclin D1 (green) in the testes from Tg (#26) mice, showing expanded SV Sertoli cells. (B) Transmission electron microscopy (TEM) analysis of a Tg (#26) testis at the RT-SV boundary (red arrow), revealing abnormal features of SV Sertoli cells, such as protruded SV (pSV) structures extending into the RT lumen (upper left in B), multiple layers of cell processes containing microtubule bundles (blue asterisks), and open extracellular spaces in the basal region of the RT-SV boundary (red asterisk). Lower panels (B’–B’’’) are magnified views of the upper right panel in B. A vertical dashed line at the center marks the region hidden by the grid (grid spacing is 10 μm). BL, multilayered SV-typical basal lamina; RT, rete testis; SV, Sertoli valve. Scale bars: 50 μm (A), 10 μm (B), 2 μm (B’–B’’’).

In this Tg line, no appreciable alterations in the RT-SV region are observed during the immature stages up to 1 week after birth. However, following sexual maturation, immunostaining for acetylated tubulin and Cyclin D1 clearly marks an expanded SV region at the terminal segments of the seminiferous tubules (Fig. 2A). This expansion was further confirmed by electron microscopy, which revealed multiple cytoplasmic layers containing bundled microtubules extending distally from Sertoli cells into the rete testis (RT) lumen (blue asterisks, Fig. 2B). At the boundary between the SV and RT regions, SV Sertoli cells were supported by a multilayered basal lamina, consistent with previous reports (Fig. S3A; Tainosho et al., 2011; Uchida et al., 2022). Notably, in the Tg testes, the SV Sertoli cells displayed vacuole-like intercellular spaces in the basal portion of the terminal segment, and some appeared detached from the basal lamina (red asterisks, Fig. 2B). Furthermore, these cells were often found protruding into the RT lumen (designated as “pSV” in Fig. 2A, 2B, S3B), where necrotic or damaged cells were observed at the distal tips of the protrusions. These findings suggest that the protruding SV Sertoli cells may undergo necrosis and be sloughed off into the RT lumen.

Partial rescue of defective spermatogenesis and infertility in RT-specific Sox17 cKO mice by AMH-Sox17 Tg

In RT-specific Sox17 conditional knockout (cKO) testes, deformation of the Sertoli valve (SV) leads to backflow of rete testis (RT) fluid into the seminiferous tubules, resulting in defective spermiogenesis and male infertility (Uchida et al., 2022). In contrast, AMH-Sox17 Tg mice exhibit excessive SV formation due to ectopic expression of SOX17 in SV Sertoli cells; however the AMH-Sox17 transgene is not activated in the RT epithelium. To evaluate whether this ectopic expression could rescue the SV defects in the RT-specific Sox17 cKO background, we examined the testis phenotype of double mutant mice (Sox17 cKO; AMH-Sox17 Tg, referred to as cKO;Tg) (Fig. 3). By crossing Tg; Sox17f/f males with SF1-Cre;Sox17f/ffemales, we generated four genotypes for comparison: control (Sox17f/f), cKO (SF1-Cre;Sox17f/f), Tg (Tg; Sox17f/f), and cKO;Tg (Sf1-Cre;Sox17f/f; Tg).

Morphometric analysis of spermatogenic quality in control (Sox17f/f), cKO (Sf1-Cre;Sox17f/f), Tg (Tg; Sox17f/f), and cKO;Tg (Sf1-Cre;Sox17f/f; Tg) testes

(A) H&E-stained sections of the epididymis and testes from control, cKO, Tg, and cKO;Tg males. Two representative testis images from both fertile and infertile cKO;Tg males are also presented. The lower panels present higher magnification images of damaged seminiferous epithelia. (B) Evaluation of spermatogenesis in cross-sections of convoluted seminiferous tubules (STs). The evaluation score (maximum 10 points) was calculated as the sum of the seminiferous tubule structural integrity score (Group A; 7 points) and the most advanced germ cell type score (Group B; 3 points). The two graphs on the right show the evaluation scores for each genotype (left; mean ± SD; *p<0.05; **p<0.01; one-way ANOVA) and the percentage of seminiferous tubules containing the most advanced germ cells (right). Defective spermatogenesis observed in cKO testes was rescued in the cKO;Tg double mutant testes. Scale bars: 50 μm (top two rows); 10 μm (bottom row) in A.

As expected, control and Tg males showed normal spermatogenesis and the presence of spermatozoa in the epididymis. In contrast, cKO males exhibited abnormal spermiogenesis with a complete absence of epididymal spermatozoa (Fig. 3A), consistent with previous findings (Uchida et al., 2022). Interestingly, cKO;Tg double mutants showed improved spermatogenesis, with the presence of epididymal spermatozoa and a partial restoration of fertility compared to cKO males. Quantitative assessment of spermatogenesis using a 10-point scoring system (7 points for seminiferous tubule integrity + 3 points for presence of advanced germ cells) confirmed a significant improvement in cKO;Tg males (Fig. 3B) related to cKO males. This was further supported by an increased ratio of elongated spermatids in these cKO;Tg testes. Notably, epididymal spermatozoa were detected in 2 of 7 cKO;Tg individuals at 8 weeks of age. Furthermore, to assess fertility, prolonged cohabitation mating trials were performed using a separate group of four cKO;Tg males. Two of the four males successfully sired offspring, although these animals were not the same as the cohorts used for the initial and quantitative evaluation of spermatogenesis.

These results demonstrate that ectopic SOX17 expression in Sertoli cells partially rescues the structural and functional defects of the SV caused by the absence of SOX17 in the RT, thereby improving spermatogenesis and restoring fertility in otherwise infertile males.

Ectopic SOX17 expression in Sertoli cells rescues SV deformation in RT-specific Sox17 cKO testes

In RT-specific Sox17 conditional knockout (cKO) testes, disruption of the SV formation leads to defective spermiogenesis and male infertility (Uchida et al., 2022), which could be partially rescued by the ectopic SOX17 expression in Sertoli cells, as described above (Fig. 3). To assess whether such ectopic SOX17 expression could rescue the SV defects caused by Sox17 deficiency in the RT, we examined the RT-SV phenotypes of Sox17 cKO;AMH-Sox17 Tg double mutant (cKO;Tg) mice (Fig. 4).

Partial normalization of the expanded SV structure in AMH-Sox17 Tg testes by cKO;Tg double mutants

(A) Immunostaining for SOX9 and SOX17 in the RT–SV–ST region of proximal testes in control, Tg, and cKO;Tg mice. Ectopic SOX17 expression is observed in Sertoli cells of Tg testes, whereas SOX17 expression is completely absent in RT cells of the cKO;Tg double mutants. The expanded region of SOX17/SOX9 double-positive SV Sertoli cells (yellow) seen in Tg testes appears to be reduced in cKO;Tg double mutants. (B) Immunostaining for SOX9 (green) and acetylated tubulin (SV marker; red) or E-cadherin (RT marker; red) demonstrates the restoration of proper SV formation in the RT–SV region of the cKO;Tg double mutants. Additional panels show immunostaining for SOX17 (red)/SOX9 (green) and HSP70 (elongated spermatid marker; red)/SCP3 (spermatocyte marker; green) in ST regions (left panels). (C) Quantitaive data of the nuclear density of Sertoli cells in the RT-SV region (mean ± SD; *p<0.05; **p<0.01; one-way ANOVA). RT, rete testis; SV, Sertoli valve. Scale bars: 200 μm (top row) and 50 μm (bottom row) in A and B.

Double immunostaining for SOX17 and SOX9 confirmed that in cKO;Tg testes, SOX17 expression was restricted to Sertoli cells in the terminal segments of the seminiferous tubules, and was absent from the RT epithelium—opposite to the expression pattern seen in control testes, where SOX17 was confined to RT epithelial cells (Fig. 4A). In Tg testes, both RT epithelial cells and SV Sertoli cells were SOX17-positive. An expanded region of SOX17/SOX9 double-positive SV Sertoli cells was evident in Tg testes, while this region appeared to be reduced in cKO;Tg testes, suggesting partial normalization of the SV region (yellow area, Fig. 4A).

Furthermore, acetylated tubulin-positive valve structures were properly formed in cKO;Tg testes, and normal spermatogenesis including SCP3-positive spermatocyte and Hsp70-positive elongated spermatids was observed in the convoluted seminiferous tubules (Fig. 4B). Quantitative analysis of SV Sertoli cell density revealed that the number of SOX9-positive nuclei in the RT-SV region was significantly increased in Sox17 Tg testes and significantly decreased in Sox17 cKO testes compared to control testes (Fig. 4C). Interestingly, in the cKO;Tg double mutant testes, SV Sertoli cell density was significantly higher than that in cKO testes, showing an intermediate level between Tg and cKO testes, comparable to that of controls (Fig. 4C). These findings indicate that ectopic SOX17 activity in SV Sertoli cells can rescue the structural defects of the SV in RT-specific Sox17 cKO testes, in a Sox17 dose-dependent manner.

Previous studies have shown that Rspo1 is expressed in RT epithelial cells in wild-type mice, but its expression is markedly reduced in cKO testes (Uchida et al., 2022). Along with the known expression of Wnt4 in SV Sertoli cells, we finally examined the expression patterns of Rspo1 and Wnt4 in the RT-SV region across different genotypes using RNAscope in situ hybridization (Fig. 5). Fixation, sectioning, and in situ hybridization were performed under identical conditions on serial sections. Expression profiles of RT-specific Rspo1 and SV-specific Wnt4 were generally comparable among control, Tg, cKO, and cKO;Tg mice (Fig. 5A), which confirm no ectopic Rspo1 expression in SOX17+ SV Sertoli cells. In Tg testes, Wnt4 expression appeared to be enhanced in the SV region, consistent with the expanded SV phenotype in these mice (Fig. 2). Notably, RT-specific Rspo1 expression was increased in Tg testes and slightly elevated in cKO;Tg mice relative to cKO mice, suggesting a potential paracrine mechanism, in which ectopic SOX17 activity in Sertoli cells might influence Rspo1 expression in adjacent RT epithelial cells.

Enhanced Rspo1 and Wnt4 expression in the RT-SV boundary of the testes in AMH-Sox17 Tg mice

RNA in situ hybridization showing the Rspo1 and Wnt4 expression profiles of RT-SV boundary in four genotypes: control (Sox17f/f), cKO (Sf1-Cre;Sox17f/f), Tg (Tg; Sox17f/f), and cKO;Tg (Sf1-Cre;Sox17f/f; Tg) testes. RT, rete testis; SV, Sertoli valve. Scale bar, 50μm.

Discussion

A previous study using RT-specific Sox17 cKO mice demonstrated that Sox17 expression is critical for Sertoli valve (SV) formation and proper spermiogenesis. Deletion of Sox17 in the rete testis (RT) disrupted the specification of modified Sertoli cells that form the SV, leading to backflow of RT fluid and abnormal detachment of immature spermatids within the seminiferous tubules (Uchida et al., 2022; Fig. 6; Fig. S1A). These findings established the essential role of SOX17⁺ RT epithelia in SV formation and clarified the physiological function of the SV in modulating fluid dynamics within the seminiferous tubules in vivo.

Schematic representation of valve structure size regulation by the AMH–Sox17 Tg mouse line

RT-specific Sox17 cKO mice exhibit deformed Sertoli valve (SV) structures, resulting in defective spermiogenesis and male infertility (Fig. 3; Uchida et al., 2022). Ectopic expression of SOX17 in Sertoli cells induces excessive SV formation in the terminal segments of the seminiferous tubule (ST), along with elevated Rspo1 expression in the rete testis (RT). Notably, this ectopic SOX17 expression partially rescues epididymal sperm production and restores fertility in cKO;Tg males. These Sox17 Tg mice, displaying SV hyperplasia and its functional rescue, provide valuable models for investigating paracrine signaling mechanisms within the RT-SV region.

In the present study, we demonstrated that ectopic expression of SOX17 in Sertoli cells induces excessive SV formation in the seminiferous tubules, along with increased Rspo1 expression in the RT. Moreover, this ectopic SOX17 expression partially rescued epididymal sperm production and restored fertility in cKO;Tg males (Fig. 6). These results suggest that SOX17 activity in the RT can be, at least in part, functionally substituted by its ectopic expression in Sertoli cells. The observed fertility recovery in cKO;Tg males, even in the absence of SOX17 in RT cells, implies that the primary role of SOX17 in the RT is restricted to SV formation in the terminal segments and is not essential for sperm transport and maturation in the RT-to-epididymis pathway in mice.

Ectopic SOX17 expression in Sertoli cells may mimic paracrine signaling normally mediated by SOX17⁺ RT epithelia in the SV region. One possible explanation is that RT epithelial cells and Sertoli cells share lineage similarities (Malolina and Kulibin, 2019; Uchida et al., 2022; Sato et al., 2026). Several key transcription factors of Sertoli cells — including WT1, NR5A1, SOX9, GATA4, and, in part, DMRT1— are also expressed in the RT epithelia (Aiyama et al., 2015; Malolina and Kulibin, 2019; Kulibin and Malolina, 2020; Uchida et al., 2022; Malolina et al., 2023). During sexual maturation, ectopic SOX17 expression in Sertoli cells may induce paracrine and/or autocrine signaling programs resembling those of RT cells, thereby promoting hyperplasia in Tg testes and facilitating SV recovery in cKO;Tg testes.

The Sox17 transgene is ectopically activated in Sertoli cells not only near the SV but also throughout the testis, whereas the SV phenotype is restricted to the RT-adjacent region of Tg testes (Figs. 1C, 2A, 4A). One possible explanation for this spatial discrepancy is that SOX17 may act on RT-lineage Sertoli-like cells located in the terminal segments adjacent to the RT. A recent study by Mayère et al. (2022) reported that PAX8⁺ RT progenitor cells, referred to as early supporting-like cells, contribute to the proximal region of the seminiferous tubules, suggesting heterogeneity in the origin of Sertoli cells near the RT during embryogenesis, even though PAX8 expression in fetal Sertoli cells is downregulated shortly after testis differentiation. It is therefore possible that ectopic SOX17 expression in these rare supporting-like cells may induce paracrine signaling similar to that of RT epithelia, thereby leading to SV phenotypes in Sertoli cells near the RT in Tg mice.

Otherwise, SOX17 has been shown to directly modulate genes encoding cell adhesion and extracellular matrix components, and to broadly influence basement membrane-associated gene networks during tubular morphogenesis (Niimi et al., 2004; Niakan et al., 2010; Walters et al., 2023). SOX17 may enhance the responsiveness of SOX17⁺ Sertoli cells to RT-derived morphogenetic signals through the establishment of an RT-like extracellular matrix microenvironment, thereby promoting spatially restricted SV formation near the RT during postnatal maturation. Indeed, RT epithelia produce high levels of growth factors and cytokines, including FGFs, TGFs, and WNT signals (Imura-Kishi et al., 2020; Uchida et al., 2022).

Interestingly, Rspo1 and Wnt4 signaling is activated in the RT-SV region (Uchida et al., 2022; this study). In Wnt4-null embryos, the RT region is underdeveloped at the seminiferous tubule–RT junction, leading to a failure in connecting the RT and seminiferous tubules, similar to the rete ovary (RO) phenotype (Mayère et al., 2022). These findings suggest that Wnt4–Rspo1 signaling is essential for proper connection of the terminal segment of seminiferous tubules to the RT and efferent ducts, which is critical for the integrity of the male reproductive tract. Our in situ hybridization analyses showed enhanced Rspo1 and Wnt4 expression driven by SOX17⁺ Sertoli cells in Tg mice (Fig. 4), which likely contributed to the rescue of SV formation observed in cKO;Tg mice (Fig. 5). The presence of protruding SV cells expressing Cyclin D1 supports the idea that SOX17-mediated paracrine activity may involve localized Rspo1-Wnt4 signaling during SV formation. Considering the potential involvement of other signaling pathways such as FGFs and TGFs, further studies are warranted to delineate the specific roles of individual paracrine factors in the RT–SV region prior to SV formation. Moreover, the present study raises several important questions regarding how SV Sertoli cells adjacent to the RT region may influence the total number of Sertoli cells per testis, and how RT-lineage Sertoli-like cells contribute to the support of spermatogenesis in the vicinity of the SV. Addressing these issues will be essential to determine how RT-derived signaling, fluid dynamics, cellular composition, lineage heterogeneity, and RT architecture collectively regulate SV formation at the RT–SV interface.

We acknowledge several limitations in the present study. First, quantitative molecular analysis of the SV region remains technically challenging. The Sertoli valve is a very small transitional structure, with approximately 20 sites per mouse testis, and is confined to a narrow region within ∼100 μm from the terminal end of the rete testis-associated seminiferous tubules (Aiyama et al., 2015; Nakata et al., 2015; Imura-Kishi et al., 2020). Because of this limited spatial extent, selective isolation of SV tissue without contamination from adjacent regions, as well as acquisition of sufficient material for quantitative analyses such as qPCR, is currently difficult (Imura-Kishi et al., 2020). In addition, the highly interdigitated architecture of Sertoli cells, which are tightly connected via specialized junctional complexes, further complicates their dissociation into intact single-cell populations after sexual maturation. For these technical reasons, SV-associated molecular changes in this study had to be primarily evaluated using histological and immunohistochemical approaches. Second, while our previous and present studies support a role for SOX17 in regulating SV formation, the direct downstream target genes and molecular mechanisms remain unidentified, despite extensive scRNA-seq analyses of SV and RT cells in RT-specific Sox17 cKO testes (Uchida et al., 2022). Using this Sox17 Tg mouse line, future comprehensive transcriptomic approaches (e.g., bulk RNA-seq and spatial transcriptomics of the SOX17+ Sertoli cells) will make it possible to identify candidate downstream effectors and further validate the proposed model in which SOX17 regulates SV formation via paracrine signaling mechanisms.

In conclusion, by generating a novel AMH-Sox17 Tg mouse line, we demonstrate that the size of the SV structure in the terminal segment of the seminiferous tubules can be experimentally manipulated. Furthermore, SV formation was successfully rescued in Sox17 cKO testes by introducing the AMH-Sox17 transgene into RT-specific Sox17 cKO mice. These genetically engineered models — exhibiting SV overdevelopment or restoration — provide valuable tools for investigating paracrine signaling and elucidating the biological significance of the SV and RT in the male reproductive system of mammals.

Materials and Methods

Animals

AMH-Sox17 Tg mice were generated by pronuclear injection of the plasmid-based transgene into the C57BL6 background, as described previously (Shinomura et al., 2014). The transgene was prepared by inserting the entire coding region of the Sox17 gene (Kanai-Azuma et al., 2002) downstream of the hAMH promoter-driven cassette (Fig. 1A). We obtained three independent F0 Tg males (lines #13, 26, and 27). Among the three founders, only line #26 was successfully established as a breeding line while maintaining spermatogenic capacity and was therefore used for most analyses in this study. Immunohistochemical analysis of Tg26 testes revealed mosaic, heterogeneous SOX17 expression among individual Sertoli cells. The #27 F0 male was infertile and exhibited extensive detachment of SV-like Sertoli cells that ectopically and uniformly expressed SOX17. Consequently, this #27 line could not be established as a breeding line. The RT-specific Sox17 cKO mice (i.e., SF1-Cre;Sox17flox/flox mice; Dhillon et al., 2006; Kim et al., 2007) were produced from the Jackson Laboratory (Uchida et al., 2022). The SF1-Cre:Sox17flox/flox;AMH-Sox17 Tg double mutant (cKO;Tg), AMH-Sox17 Tg;Sox17 flox/flox (Tg), SF1-Cre;Sox17flox/flox (cKO) and Sox17 flox/flox (control) mice were mainly generated by crossing the AMH-Sox17 Tg; Sox17 flox/flox males with SF1-Cre:Sox17 flox/flox females. These virgin mutant males at 4 and 8 weeks of age were used for the subsequent phenotypic analyses. Note that a separate cohort was utilized for this mating trials, distinct from the animals subjected to histopathological analyses of spermatogenesis. To maximize the analysis of the anatomically minute RT–SV region, the testes and epididymides from these virgin males were dissected into the proximal quarter containing the RT–SV region (for SV analysis) and the remaining tissue blocks for each animal (Figure S4). Consequently, due to these technical requirements, tissue allocation strategy, and independent evaluation of distinct histological parameters, the exact numbers of animals analyzed differed across individual assays.

All animal experiments were performed following the Guidelines for Animal Use and Experimentation at the University of Tokyo, and all experimental procedures performed herein were approved by the Institutional Animal Care and Use Committee, Graduate School of Agricultural and Life Sciences, University of Tokyo (approval IDs: P13-762, P13-764, P22-147). The PCR list for the genotyping was shown in Table S1.

Histology and immunohistochemistry

Testes were fixed in 4% PFA at 4°C overnight, dehydrated through an ethanol series, and embedded in paraffin. For the quantitative analysis of the RT-SV region, testes were transected after fixation and surface dye marking at approximately one-quarter of their length from the rete testis (RT) side. The quarter containing the RT and the proximal seminiferous tubules was used for analysis of the Sertoli valve (SV) region, whereas the remaining three-quarters were used to evaluate spermatogenesis. During the dehydration process, the RT-containing tissue block, whose cutting surface was pre-labeled with hematoxylin, was placed and gently pressed beneath a coverslip (Fig. S4), flattening the specimen and allowing the seminiferous tubules to spread radially from the RT while orienting the RT–SV–seminiferous tubule (ST) axis nearly parallel to the sectioning plane. The tissue was subsequently embedded in paraffin while maintaining this orientation. Serial sections were cut parallel to the flattened surface, enabling frequent acquisition of sagittal RT–SV–ST profiles (Fig. 4A; S4).

The paraffin sections (4 μm thickness) were subjected to hematoxylin and eosin (H&E) staining, PNA lectin (acrosome) staining and immunohistochemistry. For immunohistochemistry, the sections were incubated overnight at 4°C with the primary antibodies listed in Table S2. Immunoreactions were visualized using Alexa fluor-488/555/594/680-conjugated secondary antibodies (Abcam). Samples were analyzed by fluorescence microscopy (BX51N-34-FL-2) or Leica TCS SP8 confocal laser microscopy. Testes from at least four animals in each group were examined, and tissue sections from cKO, Tg and cKO;Tg and their littermate controls were processed in parallel to ensure the reproducibility of the results.

Electron microscopy

Testes were trimmed and fixed in 2.5% glutaraldehyde in 0.1 M sodium cacodylate buffer (pH 7.4) overnight, post-fixed with 0.5% OsO4 suspended in 0.1 M phosphate buffer (pH 7.3) for 30 min, and dehydrated through a graded ethanol series. After passing through propylene oxide, the tissues were embedded in Epon 812. Ultrathin sections were cut, stained with uranyl acetate and lead citrate, and observed by TEM (JEM-1010; JEOL).

RNA in situ hybridization

Testes were fixed in 10% formalin, dehydrated through an ethanol series, and embedded in paraffin. The paraffin sections (4 μm thickness) were processed for RNA in situ detection using the RNAscope 2.5HDAssay Red Manual (ACDBio) as described previously (Imura-Kishi et al., 2021; Uchida et al., 2022). The probes used in this study are listed in Table S3.

Evaluation of Spermatogenesis score

To evaluate spermatogenesis in mouse testes, we performed the fluorescence staining with PNA lectin (a marker for acrosome; Nakata et al., 2015) on tissue sections (Control: n = 4; cKO: n = 4; Tg: n = 5; cKO;Tg: n = 5 males). Fluorescent images of entire testis sections were acquired using a confocal microscope (TCS SP8 Leica Microsystems) with the Tiling function in LAS X software. A two-axis scoring system was applied: Group A categorized the morphological condition of the seminiferous tubules, and Group B assessed the most advanced germ cell stage present within each tubule. Scores were assigned as maximun 10 points as follows: Group A evaluated the morphological condition of the seminiferous tubules, while Group B assessed the most advanced germ cell stage present in each tubule. Scores were assigned with a maximum of 10 points (Group A [tubule condition]: normal [7 points]; disrupted spermatocyte stage [5 points]; partial cell dropout or loss [3 points]; Sertoli cells only [0 points]) plus Group B [most advanced germ cell type]: elongated spermatids [3 points]; round spermatids [2 points]; spermatocytes [1 point]; spermatogonia only [0 points]) (see Fig. 3B).

Quantitative morphometric analyses

The number of Sertoli cells in the RT-SV region was quantified using the RT-SV-ST spreading sections immunostained for SOX9, E-cadherin, and DAPI. Only sagittal sections showing a continuous RT–SV–ST axis and sectioning the SV approximately through its mid-sagittal plane were included in the quantitative analysis. Based on the distribution of spermatogonial stem cells, the ∼100 μm region extending from the RT boundary along the seminiferous tubule toward the ST side was defined as the SV region (Aiyama et al., 2015). The SV region was operationally defined as comprising (i) the terminal 100 μm segment of the seminiferous tubule immediately adjacent to the rete testis (RT) and (ii) the protruded SV extending into the RT lumen (Fig. 4C; S4). All SOX9-positive/E-cadherin-negative Sertoli cell nuclei within this region were counted.

Statistical analysis

For comparisons among four genotype groups, one-way ANOVA followed by Tukey’s multiple comparisons test was performed. Quantitative data are presented as means ± standard deviation (SD) and shown as bar graphs with dot plots. Statistical analysis was performed using GraphPad Prism (v.10.4.2; GraphPad Software, San Diego, CA, USA). P < 0.05 was considered indicative of significance, and levels of significance are shown as *P < 0.05 and **P < 0.01. In addition, single comparisons between two groups, as shown in Fig. S2A, were were analyzed using a two-tailed Student’s t-test (data are presented as means ± standard error of the mean).